Early Observations of the Type Ia Supernova \abc:\\ A Case of Interaction with Nearby, Unbound Material and/or Strong Ejecta Mixing

### 2018-05-22 doc/ms.tex
Add models section

Early-phase observations of SNe Ia from a white dwarf (WD)$+$nondegenerate
binary may detect excess emission, relative to most SNe Ia, due to the
collision of the SN ejecta with the nondegenerate companion
\citep{1973ApJ...186.1007W,2010ApJ...708.1025K}. This excess emission was
first detected in iPTF\,14atg \citep{2015Natur.521..328C}, a low-velocity SN
Ia with a significant and declining ultraviolet (UV) pulse detected within a
few days of the SN explosion. This UV pulse is best interpreted as an SN
ejecta-companion collision (but see also
\citealt{2016MNRAS.459.4428K,2017MNRAS.472.2787N}). While such emission
requires a favorable geometric alignment and is only expected in
$\lesssim$10\% of SNe Ia \citep{2010ApJ...708.1025K}, many studies have
searched for signatures of an ejecta-companion interaction, typically
resulting in nondetections (e.g.,
\citealt{2010ApJ...722.1691H,2011ApJ...741...20B,2012ApJ...744...38F,
2012ApJ...744L..17B,2015Natur.521..332O,
2013ApJ...778L..15Z,2015ApJ...799..106G,2016ApJ...826..144S,
2015ApJS..221...22I}). Possible exceptions include SN\,2012cg, which
exhibited excess blue emission in its early-phase light curve
(\citealt{2016ApJ...820...92M}; though for an interpretation that does not
invoke ejecta-companion interaction, see \citealt{2016arXiv161007601S}), and
SN\,2017cbv, which shows a clearly resolved ``bump'' in the early $UBg$
light curves \citep{2017ApJ...845L..11H}.

@@

Roughly 16\,days prior to $T_{B,_\mathrm{max}}$, the $(B - V)_0$ color of
\abc\ is $\sim$0.5\,mag bluer than SN\,2011fe. Like \abc, SN\,2012cg
\citep{2016ApJ...820...92M} and SN\,2017cbv \citep{2017ApJ...845L..11H}
exhibit $(B - V)_0$ colors that are significantly bluer than SN\,2011fe at
very early epochs. While there are many factors that contribute to the early
optical colors of SNe Ia (see \S\ref{sec:lc_energy} below), early blue
colors are often interpreted as a hallmark of interaction between the SN
ejecta and a binary companion. Nevertheless, it is interesting to note that
despite the blue optical colors, the $\mathrm{UV} - \mathrm{optical}$ colors
of \abc, SN\,2012cg, and SN\,2017cbv are significantly redder at these early
epochs than the $\mathrm{UV} - \mathrm{optical}$ colors of iPTF\,14atg
\citep{2015Natur.521..328C}, the most likely candidate for SN
ejecta-companion interaction.

@@

Relative to the nearby, normal SN\,2011fe, we have identified several distinct
characteristics of the early evolution of \abc, including: (i) a near-linear
photometric rise, (ii) a qualitatively short, or possibly absent, dark phase,
assuming $v_\mathrm{ph} \propto t^{-0.22}$, (iii) the presence of strong
\ion{C}{2} absorption, and (iv) blue and nearly constant $(B - V)_0$
color in the week after explosion. While most SNe Ia are powered purely by
radioactive decay, the observed radiation shortly after explosion can also
include contributions from SN shock cooling or the collision of the SN ejecta
with a nondegenerate companion or nearby, unbound material. Here we
consider these scenarios as possible explanations for the early behavior of
\abc.

@@

The shock breakout of a SN Ia lasts for a fraction of a second due to compact
size of the exploding star. Emission from the subsequent cooling phase may
last for several days, however (e.g., \citealt{2010ApJ...708..598P}).
Following the analysis of \citet{2012ApJ...744L..17B} for SN\,2011fe, we
compare the early-phase $g_\mathrm{PTF}$ light curve of \abc\ with two shock
cooling models \citep{2011ApJ...728...63R, 2010ApJ...708..598P}. From this
analysis, we constrain the \abc\ progenitor radius to be $<1\,\sr$. Our
observations of \abc\ cannot place tight constraints on the size of its
progenitor. Indeed, for a typical WD radius, such as that inferred for
SN\,2011fe ($\lesssim 0.02$--$0.04\,\sr$; \citealt{2012ApJ...744L..17B,
2014ApJ...784...85P}), the expected emission from shock cooling is
$\sim$2\,mag fainter than the P48 $g_\mathrm{PTF}$ detection limit at this
distance. Thus, we conclude that shock cooling does not contribute to the
early emission detected from \abc.

\subsection{SN-Companion collision}
\label{sec:companion}

The detection of emission from the collision of the SN ejecta with a
nondegenerate companion requires a favorable orbital alignment relative to
the line of sight. Thus, from geometric considerations alone the probability
of detecting ejecta-companion interaction is low, $\sim$10\%.
\citet{2010ApJ...708.1025K} calculates that the collision of SN ejecta with a
companion generates thermal emission with a spectrum that peaks in the UV. The
resulting \textit{g}-band emission is expected to be weak.

To examine the possibility of a SN-companion signature in the early light
curve of \abc, we employ the \citet{2010ApJ...708.1025K} model and assume
canonical values for the ejecta mass, $1.4\,\sm$, expansion velocity,
$10^{4}\,\textrm{km}\,\textrm{s}^{-1}$, and a constant opacity,
$0.2\,\textrm{cm}^2\,\textrm{g}^{-1}$. We calculate the expected
$g_\mathrm{PTF}$ brightness of an ejecta-companion collision at the distance
of \abc\ behind a total reddening of $E(B-V) = 0.08 \, \mathrm{mag}$ using the
parameterized equations in \citet{2012ApJ...749...18B}. If we assume the
binary is aligned with the optimal orientation relative to the line of sight,
a binary separation of $a \approx 3 \times 10^{11}\, \mathrm{cm}$ is needed to
explain the initial detection of \abc, as shown in
Figure~\ref{fig:SN-companion}. The minimum binary separation capable of
explaining the observed brightness at the epoch of discovery is $a \approx
10^{11} \, \mathrm{cm}$. Figure~\ref{fig:SN-companion} shows that such models
peak at $g_\mathrm{PTF} \approx 21.5 \, \mathrm{mag}$, provided that
$t_\mathrm{exp} \approx t_0 - 0.3 \, \mathrm{d}$. These models do not,
however, match the $g_\mathrm{PTF}$ evolution for $t > t_0 + 0.5 \,
\mathrm{d}$ (though it is possible that the SN photosphere dominates the
companion-interaction signature at this phase). While they are otherwise
compatible with the observations, we do not favor the above models as the
explanation for the early flux from \abc\ because they do not explain the
2\,mag rise in the $\sim$24 hr after discovery.

\begin{figure}[!thb]
  \centering
  \includegraphics[width=3.35in]{companion_lc.pdf}
  \caption{Comparison of SN ejecta-companion interaction models 
  with early observations of \abc. $g_\mathrm{PTF}$ 
  detections and 3$\sigma$ upper limits are shown 
  as green circles and downward arrows, respectively. 
  \textit{Swift}/$UVW1$ observations are shown as magenta hexagons.
  The dashed, solid, and dash-dot lines show the expected flux 
  for companion-interaction models in the $g_\mathrm{PTF}$ (green) 
  and $UVW1$ (magenta) filters. The models have been adjusted to 
  account for the distance and reddening towards \abc. Each 
  model features a different companion semi-major axis, $a$, and 
  time of explosion, $t_\mathrm{exp}$, as labeled in the 
  legend. While models with $a \gtrsim 10^{12} \, \mathrm{cm}$ can 
  explain the early optical rise, they greatly 
  overpredict the UV flux. Models with $a \approx 10^{11} \, 
  \mathrm{cm}$ can explain the initial detection \abc, but they 
  fail to replicate the $\sim$2\,mag rise in the $\sim$24\,hr 
  after explosion.
  }
  \label{fig:SN-companion}
\end{figure}

Figure~\ref{fig:SN-companion} additionally shows that a companion at $a
\approx 18 \times 10^{11} \, \mathrm{cm}$ provides a good match to the initial
optical rise, if $t_\mathrm{exp} \approx t_0 + 0.1 \; \mathrm{d}$. Models with
$a \gtrsim 10^{12} \, \mathrm{cm}$, which can explain the initial
$g_\mathrm{PTF}$ rise, significantly overpredict the observed UV flux,
however. There is no choice of $a$ capable of replicating the early rise of
\abc\ without also overpredicting the observed UV flux.

The challenges associated with each of the previously considered models lead
us to conclude that the early evolution of \abc\ cannot be explained via
ejecta-companion interaction. We cannot, however, exclude the presence of a
red-giant, or other nondegenerate, companion as our calculations have assumed
that the binary is aligned with the optimal geometry relative to the line of
sight. If the geometry is not favorable, then it is possible that signatures
from interaction with a companion are not visible.

@@

In \citet{2017MNRAS.472.2787N} the early photometric evolution of Type
Ia SNe is explored via a variety of explosion models and detailed
radiative-transfer calculations. Specifically, \citet{2017MNRAS.472.2787N}
examine two Chandrasekhar-mass ($M_\mathrm{Ch}$) explosions and compare their
evolution to sub-Chandrasekhar detonations and pure deflagrations. The
$M_\mathrm{Ch}$ explosions include the ``W7'' carbon-deflagration model of
\citet{1984ApJ...286..644N} and the ``N100'' delayed-detonation model from
\citet{2013MNRAS.429.1156S}. The sub-Chandrasekhar models include: a violent
WD-WD merger, which triggers a carbon detonation in the more massive WD, a
centrally ignited sub-Chandrasehkar detonation, and a sub-Chandrasehkar double
detonation explosion, in which a He-surface-layer detonation triggers a carbon
detonation in the core. \citet{2017MNRAS.472.2787N} note that, of these last
two sub-Chandrasehkar models, the later provides the more realistic scenario.
Finally, \citeauthor{2017MNRAS.472.2787N} also examine the ``N5def'' and
``N1600Cdef'' pure deflagration explosions from \citet{2014MNRAS.438.1762F}.
The ``N5def'' model is particularly unique in that the explosion does not
fully unbind the WD, meaning, unlike typical Type Ia SNe, a remnant remains.

\begin{figure}[]
  \centering
  \includegraphics[width=3.35in]{compare_Noebauer.pdf}
  \caption{
  %
  Comparison of the models from \citet{2017MNRAS.472.2787N to the
  \abc\ $V$-band light curve (\textit{top}) and the $(B-V)_0$ color curve
  (\textit{bottom}). The light curve has been corrected for the distance
  modulus to \abc\ (\S\ref{sec:host}), while both the light curve and color
  curve have been corrected for reddening. To guide the eye,
  $g_\mathrm{PTF}$ observations are shown, though we caution
  that \abc\ may exhibit significant color evolution at this phase in which
  case $g_\mathrm{PTF}$ would be a poor proxy for $V$. Each model light curve
  is translated to match the LCO observations $\sim$1.7\,d after $t_0$ given
  the uncertain time of explosion. Translational offsets are listed in the
  bottom panel legend, which shows the models in order of decreasing
  $^{56}$Ni from top to bottom, then left to right. The inset in
  the top panel shows the residuals relative to the W7 model.} }
  %
  \label{fig:noebauer} 
\end{figure}

In Figure~\ref{fig:noebauer} we compare photometric observations of
\abc\ to the models presented in
\citet{2017MNRAS.472.2787N}.\footnote{Available here:
\url{https://hesma.h-its.org/}.} Interestingly, the sub-Chandrasekhar double
detonation model (subChDoubleDet) replicates the early wiggle in the $V$-band
light curve. However, this match requires an explosion \textit{after} our
initial detection of \abc, and predicts extreme color evolution that is not
observed. Thus, the subChDoubleDet model is incompatible with the
observations. The sub-Chandrasekhar detonation (subChDet) provides a better
match to the observations, though this model is not favored as a particularly
realistic scenario (see above). Of the sub-Chandrasekhar models the violent
merger model (Merger) provides the best match to the observations including
the early rise and color evolution. In detail, however, this model does not
match the early wiggles in the light curve, has consistently redder colors
than \abc, and requires $t_\mathrm{exp} \approx 2.3$\,d prior to $t_0$. As
such, we postulate that \abc\ is not the result of a violent WD-WD merger.

For clarity, of the two pure deflagration models only N5def is shown,
however, the evolution of N1600Cdef is very similar. While the pure
deflagration models produce the bluest colors at early times, they are
underluminous at times $> t_0 + 4\,\mathrm{d}$ and already rapidly evolving
towards the red at $\sim$$t_0 + 7$ (\abc\ exhibits a nearly constant $(B-V)_0$
color for $\sim$19\,d after $t_0$). Thus, we conclude that \abc\ is not
compatible with pure deflagrations.

Of the $M_\mathrm{Ch}$ models, the W7 model better matches the
observations as the N100 model features a faster rise and higher luminosity
than what is observed. We explore delayed detonation models in further detail
below.}

@@

To model SN\,2011fe, \citet{2014MNRAS.441..532D} examined
pulsational delayed-detonation (PDD) models as an explanation for some SNe
Ia. Briefly, PDD models differ from ``standard'' delayed detonation (DD)
models in that the expansion of the WD during the initial deflagration
phase leads to the release of unbound material. Following this pulsation,
the bound material contracts eventually triggering a subsequent
detonation.\footnote{\citet{2014MNRAS.441..532D} note that the deflagration
and detonation in their PDD models are artificially triggered.} An
important consequence of this progression for PDD models is that the
unbound material expands and avoids burning, unlike DD models that
typically leave no unburnt material. This results in significantly more
carbon in the outer layers of the SN ejecta \citep{2014MNRAS.441..532D}.

\citet{2014MNRAS.441..532D} find that DD models are universally faint and red at early times, $\sim$24--48\,hr after explosion, while the PDD models exhibit a faster rise and bluer colors. Briefly, this occurs in the PDD scenario because the collision with the unbound material surrounding the WD heats the outer layers of the SN ejecta. Importantly, the PDD models are nearly indistinguishable from DD models around and post peak.

\begin{figure}[]
  \centering
  \includegraphics[width=3.35in]{compare_Dessart.pdf}
  \caption{
  %
  Same as Figure~\ref{fig:noebauer}, featuring the $g$-band light
  curve and the DD and PDD models from \citet{2014MNRAS.441..532D}. The bottom
  panel legend lists the models in order of decreasing $M_\mathrm{Ni}$ from
  top to bottom. DD models (labeled as DDC to match the nomenclature of
  \citealt{2014MNRAS.441..532D}) are shown as dash-dot lines, while PDD models
  (labeled as PDDEL) are shown as dashed lines. The inset in the top panel
  shows the residuals relative to the PDDEL7n model. The PDD models provide a
  better match to the observations.}
  %
  \label{fig:dessart} 
\end{figure}

In Figure~\ref{fig:dessart} we compare photometric observations of
\abc\ to the high-$^{56}$Ni yield DD and PDD models presented in
\citet{2014MNRAS.441..532D}.\footnote{Available here:
\url{https://www-n.oca.eu/supernova/snia/snia_ddc_pddel.html}.} DD and PDD
models with $M_\mathrm{Ni}/M_\odot \la 0.7$ and 0.5, respectively, fail to
match the early luminosity and blue colors of \abc. The PDD models provide a
better match to the observations than the DD models. The PDD models evolve
more rapidly towards blue colors and provide a better match to the $g$-band
flux in the days after explosion. Furthermore, unlike the DD models, the PDD
models exhibit strong \ion{C}{2} lines that gradually disappear in the
$\sim$week after explosion \citep{2014MNRAS.441..532D}. In detail, the early
wiggles in the \abc\ light curve are not matched by the PDD models, which
exhibit more smooth variations. Furthermore, while the focus of this study is
the early evolution of \abc, the PDD models evolve more rapidly to the red
post-peak than the observations. Nevertheless, the PDD models presented in
\citet{2014MNRAS.441..532D} provide several attractive explanations for the
unusual features in the early behavior of \abc. Small adjustments to the PDD
models (e.g., additional $^{56}$Ni mixing, which
\citeauthor{2014MNRAS.441..532D} only explore for DD models) may better match
\abc.

\subsection{Strong $^{56}Ni$ Mixing in the SN Ejecta}
\label{sec:Ni_mixing}

Having examined other possibilities we now consider whether the early evolution of \abc\ can be explained simply by invoking strong mixing in the 
SN ejecta. Strong mixing leads to a faster initial rise as well as a more rapid evolution towards blue colors.

\begin{figure}[]
  \centering
  \includegraphics[width=3.35in]{compare_Piro.pdf}
  \caption{
  %
  Same as Figure~\ref{fig:noebauer}, featuring the $g$-band light
  curve and models from \citet{2016ApJ...826...96P}. The amount of $^{56}$Ni
  mixing in the SN ejecta increases from the short-dash, light lines to the
  long-dash, dark lines. Unlike Figure~\ref{fig:dessart}, each model features
  the same $M_\mathrm{Ni}$, while the model names reflect the boxcar widths
  used to approximate the effects of mixing in the ejecta (see
  \citealt{2016ApJ...826...96P}). The top panel inset shows the residuals
  relative to the $0.25\,M_\odot$ model. The observations are best matched by
  the $0.25\,M_\odot$ model, i.e. the model with the most significant mixing.}
  %
  \label{fig:piro}
\end{figure}

>>>>

Early-phase observations of SNe Ia from a white dwarf (WD)$+$nondegenerate
binary may detect excess emission, relative to most SNe Ia, due to the
collision of the SN ejecta with the nondegenerate companion
\citep{1973ApJ...186.1007W,2010ApJ...708.1025K}. This excess emission was
first detected in iPTF\,14atg \citep{2015Natur.521..328C}, a low-velocity SN
Ia with a significant and declining ultraviolet (UV) pulse detected within a
few days of the SN explosion. This UV pulse is best interpreted as an SN
ejecta--companion collision (but see also
\citealt{2016MNRAS.459.4428K,2017MNRAS.472.2787N}). While such emission
requires a favorable geometric alignment and is only expected in
$\lesssim$10\% of SNe Ia \citep{2010ApJ...708.1025K}, many studies have
searched for signatures of an ejecta--companion interaction, typically
resulting in nondetections (e.g.,
\citealt{2010ApJ...722.1691H,2011ApJ...741...20B,2012ApJ...744...38F,
2012ApJ...744L..17B,2015Natur.521..332O,
2013ApJ...778L..15Z,2015ApJ...799..106G,2016ApJ...826..144S,
2015ApJS..221...22I}). Possible exceptions include SN\,2012cg, which
exhibited excess blue emission in its early-phase light curve
(\citealt{2016ApJ...820...92M}; though for an interpretation that does not
invoke ejecta--companion interaction, see \citealt{2016arXiv161007601S}), and
SN\,2017cbv, which shows a clearly resolved ``bump'' in the early $UBg$
light curves \citep{2017ApJ...845L..11H}.

@@

Roughly 16\,days prior to $T_{B,_\mathrm{max}}$, the $(B - V)_0$ color of
\abc\ is $\sim$0.5\,mag bluer than SN\,2011fe. Like \abc, SN\,2012cg
\citep{2016ApJ...820...92M} and SN\,2017cbv \citep{2017ApJ...845L..11H}
exhibit $(B - V)_0$ colors that are significantly bluer than SN\,2011fe at
very early epochs. While there are many factors that contribute to the early
optical colors of SNe Ia (see \S\ref{sec:lc_energy} below), early blue
colors are often interpreted as a hallmark of interaction between the SN
ejecta and a binary companion. Nevertheless, it is interesting to note that
despite the blue optical colors, the $\mathrm{UV} - \mathrm{optical}$ colors
of \abc, SN\,2012cg, and SN\,2017cbv are significantly redder at these early
epochs than the $\mathrm{UV} - \mathrm{optical}$ colors of iPTF\,14atg
\citep{2015Natur.521..328C}, the most likely candidate for SN
ejecta--companion interaction.

@@

Relative to the nearby, normal SN\,2011fe, we have identified several
distinct characteristics of the early evolution of \abc, including (i) a
near-linear photometric rise; (ii) a qualitatively short, or possibly
absent, dark phase, assuming $v_\mathrm{ph} \propto t^{-0.22}$; (iii) the
presence of strong \ion{C}{2} absorption; and (iv) blue and nearly constant
$(B - V)_0$ color in the week after explosion. While most SNe Ia are powered
purely by radioactive decay, the observed radiation shortly after explosion
can also include contributions from SN shock cooling or the collision of the
SN ejecta with a nondegenerate companion or nearby, unbound material. Here
we consider these scenarios as possible explanations for the early behavior
of \abc.

@@

The shock breakout of an SN Ia lasts for a fraction of a second due to compact
size of the exploding star. Emission from the subsequent cooling phase may
last for several days, however (e.g., \citealt{2010ApJ...708..598P}).
Following the analysis of \citet{2012ApJ...744L..17B} for SN\,2011fe, we
compare the early-phase $g_\mathrm{PTF}$ light curve of \abc\ with two shock
cooling models \citep{2010ApJ...708..598P, 2011ApJ...728...63R}. From this
analysis, we constrain the \abc\ progenitor radius to be $<1\,\sr$. Our
observations of \abc\ cannot place tight constraints on the size of its
progenitor. Indeed, for a typical WD radius, such as that inferred for
SN\,2011fe ($\lesssim 0.02$--$0.04\,\sr$; \citealt{2012ApJ...744L..17B,
2014ApJ...784...85P}), the expected emission from shock cooling is
$\sim$2\,mag fainter than the P48 $g_\mathrm{PTF}$ detection limit at this
distance. Thus, we conclude that shock cooling does not contribute to the
early emission detected from \abc.

\subsection{SN--Companion Collision}
\label{sec:companion}

The detection of emission from the collision of the SN ejecta with a
nondegenerate companion requires a favorable orbital alignment relative to
the line of sight. Thus, from geometric considerations alone the probability
of detecting ejecta--companion interaction is low, $\sim$10\%.
\citet{2010ApJ...708.1025K} calculates that the collision of SN ejecta with
a companion generates thermal emission with a spectrum that peaks in the UV.
The resulting \textit{g}-band emission is expected to be weak.

To examine the possibility of a SN--companion signature in the early light
curve of \abc, we employ the \citet{2010ApJ...708.1025K} model and assume
canonical values for the ejecta mass, $1.4\,\sm$, expansion velocity,
$10^{4}\,\textrm{km}\,\textrm{s}^{-1}$, and a constant opacity,
$0.2\,\textrm{cm}^2\,\textrm{g}^{-1}$. We calculate the expected
$g_\mathrm{PTF}$ brightness of an ejecta--companion collision at the
distance of \abc\ behind a total reddening of $E(B-V) = 0.08 \,
\mathrm{mag}$ using the parameterized equations in
\citet{2012ApJ...749...18B}. If we assume that the binary is aligned with
the optimal orientation relative to the line of sight, a binary separation
of $a \approx 3 \times 10^{11}\, \mathrm{cm}$ is needed to explain the
initial detection of \abc, as shown in Figure~\ref{fig:SN--companion}. The
minimum binary separation capable of explaining the observed brightness at
the epoch of discovery is $a \approx 10^{11} \, \mathrm{cm}$.
Figure~\ref{fig:SN--companion} shows that such models peak at
$g_\mathrm{PTF} \approx 21.5 \, \mathrm{mag}$, provided that $t_\mathrm{exp}
\approx t_0 - 0.3 \, \mathrm{days}$. These models do not, however, match the
$g_\mathrm{PTF}$ evolution for $t > t_0 + 0.5 \, \mathrm{days}$ (though it
is possible that the SN photosphere dominates the companion-interaction
signature at this phase). While they are otherwise compatible with the
observations, we do not favor the above models as the explanation for the
early flux from \abc\ because they do not explain the 2\,mag rise in the
$\sim$24 hr after discovery.

\begin{figure}[!thb]
  \centering
  \includegraphics[width=3.35in]{companion_lc.pdf}
  %
  \caption{Comparison of SN ejecta--companion interaction models with early
  observations of \abc. $g_\mathrm{PTF}$ detections and 3$\sigma$ upper
  limits are shown as green circles and downward-pointing arrows,
  respectively. \textit{Swift}/$UVW1$ observations are shown as magenta
  hexagons. The dashed, solid, and dot-dashed lines show the expected flux
  for companion-interaction models in the $g_\mathrm{PTF}$ (green) and
  $UVW1$ (magenta) filters. The models have been adjusted to account for the
  distance and reddening toward \abc. Each model features a different
  companion semi-major axis, $a$, and time of explosion, $t_\mathrm{exp}$,
  as labeled in the legend. While models with $a \gtrsim 10^{12} \,
  \mathrm{cm}$ can explain the early optical rise, they greatly overpredict
  the UV flux. Models with $a \approx 10^{11} \, \mathrm{cm}$ can explain
  the initial detection of \abc, but they fail to replicate the $\sim$2\,mag
  rise in the $\sim$24\,hr after explosion. }
  %
  \label{fig:SN--companion}
\end{figure}

Figure~\ref{fig:SN--companion} additionally shows that a companion at $a
\approx 18 \times 10^{11} \, \mathrm{cm}$ provides a good match to the
initial optical rise, if $t_\mathrm{exp} \approx t_0 + 0.1 \;
\mathrm{days}$. Models with $a \gtrsim 10^{12} \, \mathrm{cm}$, which can
explain the initial $g_\mathrm{PTF}$ rise, significantly overpredict the
observed UV flux, however. There is no choice of $a$ capable of replicating
the early rise of \abc\ without also overpredicting the observed UV flux.

The challenges associated with each of the previously considered models lead
us to conclude that the early evolution of \abc\ cannot be explained via
ejecta--companion interaction. We cannot, however, exclude the presence of a
red giant, or other nondegenerate, companion as our calculations have
assumed that the binary is aligned with the optimal geometry relative to the
line of sight. If the geometry is not favorable, then it is possible that
signatures from interaction with a companion are not visible.

@@

In \citet{2017MNRAS.472.2787N} the early photometric evolution of SNe Ia is
explored via a variety of explosion models and detailed radiative transfer
calculations. Specifically, \citet{2017MNRAS.472.2787N} examine two
Chandrasekhar-mass ($M_\mathrm{Ch}$) explosions and compare their evolution
to sub-Chandrasekhar detonations and pure deflagrations. The $M_\mathrm{Ch}$
explosions include the ``W7'' carbon-deflagration model of
\citet{1984ApJ...286..644N} and the ``N100'' delayed-detonation model from
\citet{2013MNRAS.429.1156S}. The sub-Chandrasekhar models include a violent
WD-WD merger, which triggers a carbon detonation in the more massive WD, a
centrally ignited sub-Chandrasehkar detonation, and a sub-Chandrasehkar
double-detonation explosion, in which an He-surface-layer detonation
triggers a carbon detonation in the core. \citet{2017MNRAS.472.2787N} note
that, of these last two sub-Chandrasehkar models, the latter provides the
more realistic scenario. Finally, \citeauthor{2017MNRAS.472.2787N} also
examine the ``N5def'' and ``N1600Cdef'' pure deflagration explosions from
\citet{2014MNRAS.438.1762F}. The ``N5def'' model is particularly unique in
that the explosion does not fully unbind the WD, meaning that, unlike with
typical SNe Ia, a remnant remains.

\begin{figure}[]
  \centering
  \includegraphics[width=3.35in]{compare_Noebauer.pdf}
  \caption{
  %
  Comparison of the models from \citet{2017MNRAS.472.2787N to the \abc\
  $V$-band light curve (\textit{top}) and the $(B-V)_0$ color curve
  (\textit{bottom}). The light curve has been corrected for the distance
  modulus to \abc\ (\S\ref{sec:host}), while both the light curve and color
  curve have been corrected for reddening. To guide the eye,
  $g_\mathrm{PTF}$ observations are shown, though we caution that \abc\ may
  exhibit significant color evolution at this phase, in which case
  $g_\mathrm{PTF}$ would be a poor proxy for $V$. Each model light curve is
  translated to match the LCO observations $\sim$1.7\,days after $t_0$ given
  the uncertain time of explosion. Translational offsets are listed in the
  bottom panel legend, which shows the models in order of decreasing
  $^{56}$Ni from top to bottom, then left to right. The inset in the top
  panel shows the residuals relative to the W7 model.} }
  %
  \label{fig:noebauer} 
\end{figure}

In Figure~\ref{fig:noebauer} we compare photometric observations of \abc\ to
the models presented in \citet{2017MNRAS.472.2787N}.\footnote{Available at
\url{https://hesma.h-its.org/}.} Interestingly, the sub-Chandrasekhar
double-detonation model (subChDoubleDet) replicates the early wiggle in the
$V$-band light curve. However, this match requires an explosion
\textit{after} our initial detection of \abc and predicts extreme color
evolution that is not observed. Thus, the subChDoubleDet model is
incompatible with the observations. The sub-Chandrasekhar detonation
(subChDet) provides a better match to the observations, though this model is
not favored as a particularly realistic scenario (see above). Of the
sub-Chandrasekhar models, the violent merger model (Merger) provides the
best match to the observations, including the early rise and color
evolution. In detail, however, this model does not match the early wiggles
in the light curve, has consistently redder colors than \abc, and requires
$t_\mathrm{exp} \approx 2.3$\,days prior to $t_0$. As such, we postulate
that \abc\ is not the result of a violent WD--WD merger.

For clarity, of the two pure deflagration models only N5def is shown;
however, the evolution of N1600Cdef is very similar. While the pure
deflagration models produce the bluest colors at early times, they are
underluminous at times $>t_0 + 4\,\mathrm{days}$ and already rapidly
evolving toward the red at $\sim$$t_0 + 7$ (\abc\ exhibits a nearly constant
$(B-V)_0$ color for $\sim$19\,days after $t_0$). Thus, we conclude that
\abc\ is not compatible with pure deflagrations.

Of the $M_\mathrm{Ch}$ models, the W7 model better matches the
observations, as the N100 model features a faster rise and higher luminosity
than what is observed. We explore delayed-detonation models in further detail
below.

@@

To model SN\,2011fe, \citet{2014MNRAS.441..532D} examined pulsational
delayed-detonation (PDD) models as an explanation for some SNe Ia. Briefly,
PDD models differ from ``standard'' delayed-detonation (DD) models in that
the expansion of the WD during the initial deflagration phase leads to the
release of unbound material. Following this pulsation, the bound material
contracts, eventually triggering a subsequent
detonation.\footnote{\citet{2014MNRAS.441..532D} note that the deflagration
and detonation in their PDD models are artificially triggered.} An important
consequence of this progression for PDD models is that the unbound material
expands and avoids burning, unlike DD models that typically leave no unburnt
material. This results in significantly more carbon in the outer layers of
the SN ejecta \citep{2014MNRAS.441..532D}.

\citet{2014MNRAS.441..532D} find that DD models are universally faint and
red at early times, $\sim$24--48\,hr after explosion, while the PDD models
exhibit a faster rise and bluer colors. Briefly, this occurs in the PDD
scenario because the collision with the unbound material surrounding the WD
heats the outer layers of the SN ejecta. Importantly, the PDD models are
nearly indistinguishable from DD models around peak and post-peak.

\begin{figure}[]
  \centering
  \includegraphics[width=3.35in]{compare_Dessart.pdf}
  \caption{
  %
  Same as Figure~\ref{fig:noebauer}, but featuring the $g$-band light curve
  and the DD and PDD models from \citet{2014MNRAS.441..532D}. The bottom
  panel legend lists the models in order of decreasing $M_\mathrm{Ni}$ from
  top to bottom. DD models (labeled as DDC to match the nomenclature of
  \citealt{2014MNRAS.441..532D}) are shown as dot-dashed lines, while PDD
  models (labeled as PDDEL) are shown as dashed lines. The inset in the top
  panel shows the residuals relative to the PDDEL7n model. The PDD models
  provide a better match to the observations.}
  %
  \label{fig:dessart} 
\end{figure}

In Figure~\ref{fig:dessart} we compare photometric observations of \abc\ to
the high-$^{56}$Ni yield DD and PDD models presented in
\citet{2014MNRAS.441..532D}.\footnote{Available at
\url{https://www-n.oca.eu/supernova/snia/snia_ddc_pddel.html}.} DD and PDD
models with $M_\mathrm{Ni}/M_\odot \la 0.7$ and 0.5, respectively, fail to
match the early luminosity and blue colors of \abc. The PDD models provide a
better match to the observations than the DD models. The PDD models evolve
more rapidly toward blue colors and provide a better match to the $g$-band
flux in the days after explosion. Furthermore, unlike the DD models, the PDD
models exhibit strong \ion{C}{2} lines that gradually disappear in the
$\sim$1\,week after explosion \citep{2014MNRAS.441..532D}. In detail, the
early wiggles in the \abc\ light curve are not matched by the PDD models,
which exhibit more smooth variations. Furthermore, while the focus of this
study is the early evolution of \abc, the PDD models evolve more rapidly to
the red post-peak than the observations. Nevertheless, the PDD models
presented in \citet{2014MNRAS.441..532D} provide several attractive
explanations for the unusual features in the early behavior of \abc. Small
adjustments to the PDD models (e.g., additional $^{56}$Ni mixing, which
\citeauthor{2014MNRAS.441..532D} only explore for DD models) may better
match \abc.

\subsection{Strong $^{56}$Ni Mixing in the SN Ejecta}
\label{sec:Ni_mixing}

Having examined other possibilities, we now consider whether the early
evolution of \abc\ can be explained simply by invoking strong mixing in the
SN ejecta. Strong mixing leads to a faster initial rise, as well as a more
rapid evolution toward blue colors.

\begin{figure}[]
  \centering
  \includegraphics[width=3.35in]{compare_Piro.pdf}
  \caption{
  %
  Same as Figure~\ref{fig:noebauer}, but featuring the $g$-band light curve
  and models from \citet{2016ApJ...826...96P}. The amount of $^{56}$Ni
  mixing in the SN ejecta increases from the light, short-dashed lines to
  the dark, long-dashed lines. Unlike Figure~\ref{fig:dessart}, each model
  features the same $M_\mathrm{Ni}$, while the model names reflect the
  boxcar widths used to approximate the effects of mixing in the ejecta (see
  \citealt{2016ApJ...826...96P}). The top panel inset shows the residuals
  relative to the $0.25\,M_\odot$ model. The observations are best matched
  by the $0.25\,M_\odot$ model, i.e. the model with the most significant
  mixing.}
  %
  \label{fig:piro}
\end{figure}

Figure~\ref{fig:piro} compares the models from \citet{2016ApJ...826...96P}
to \abc. The \citeauthor{2016ApJ...826...96P} models employ a piston-driven
explosion to explode a single WD progenitor model. As the piston explosion
does not result in any nucleosynthesis, the distribution of $^{56}$Ni in the
ejecta must be prescribed by hand, which enables a study of the effects of
mixing on the resulting SN emission. Each model employs a fixed $0.5\,\sm$
of $^{56}$Ni that has been distributed throughout the ejecta via boxcar
averaging (see their Figure~1). The resulting light curves are synthesized
using the SuperNova Explosion Code (\texttt{SNEC};
\citealt{2015ApJ...814...63M}), as shown in Figure~\ref{fig:piro}. Broadly
speaking, the results can be summarized as follows: SNe with strong mixing
exhibit a rapid rise and quickly develop blue colors, whereas models where
the $^{56}$Ni is confined to the innermost layers of the ejecta remain faint
for days after explosion and feature a gradual color evolution from the red
to the blue. The model with the strongest mixing (dark long-dashed line in
Figure~\ref{fig:piro)} best matches the observations of \abc. This is the
only model we have found that exhibits a flat $(B-V)_0$ color evolution in
the days after explosion; however, in detail this model is too red relative
to the observations of \abc.

While the models from \citet{2016ApJ...826...96P} provide a good match to
the optical photometric evolution of \abc, they consistently overpredict the
flux in the UV. They also overpredict the photospheric velocity of \abc\ by
$\sim$2--3000$\; \mathrm{km \; s}^{-1}$. Furthermore, the simple gray
opacities in SNEC likely produce a faster rise and bluer colors than the
more detailed treatments employed in \citet{2014MNRAS.441..532D} and
\citet{2017MNRAS.472.2787N}.
